Type-Safe Modular Parsing

### 2017-01-14 paper/sections/RelatedWork.tex
All refs are fixed except Introduction

\paragraph{Extensibility} Various design patterns~\cite{}\huang{???} in multiple
languages, have been proposed over the years to address extensibility
problems, such as the Expression Problem~\cite{wadler1998expression}. However, to the best
of our knowledge, none of the existing extensibility techniques
address the problem of \emph{modularly
parsing extensible ASTs}.  There has been
a lot of work on extensible ASTs, in both functional and
object-oriented programming.  For instance, the famous ``Datatypes \`a
la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
of every two functors, which forms a binary tree structure.
Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
All of that work essentially covers how to traverse and
consume extensible ASTs.
Only in Bahr's~\cite{Bahr2011} work a brief mention of \emph{unfolds}
(which are a class of algorithms that \emph{build} ASTs) are mentioned.
However Bahr's work does not cover parsing.

@@

It is worth mentioning that Scala case classes~\cite{}\huang{???} provide a near
solution to the Expression Problem. Case classes are similar to
algebraic datatypes but, unlike algebraic datatypes, new cases can be
modularly added. The reason why case classes are not a complete
solution to the Expression Problem is that they do not enforce
exhaustiveness of pattern matching for extensible operations. In other
words run-time pattern matching errors can happen when writing extensible code with case classes. So that full static type-safety is not ensured. Nevertheless case classes are a pragmatic approach, which is
widely used in practice. Therefore, modular parsing techniques may be
of value for extensible code using case classes.  The approach we
presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
all that the users need to do is to use case classes, instead of
standard OO classes in their code.

>>>>

\paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
languages, have been proposed over the years to address extensibility
problems, such as the Expression Problem~\cite{wadler1998expression}. However, to the best
of our knowledge, none of the existing extensibility techniques
address the problem of \emph{modularly
parsing extensible ASTs}.  There has been
a lot of work on extensible ASTs, in both functional and
object-oriented programming.  For instance, the famous ``Datatypes \`a
la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
of every two functors, which forms a binary tree structure.
Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
All of that work essentially covers how to traverse and
consume extensible ASTs.
Only in Bahr's~\cite{Bahr2011} work a brief mention of \emph{unfolds}
(which are a class of algorithms that \emph{build} ASTs) are mentioned.
However Bahr's work does not cover parsing.

@@

It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
solution to the Expression Problem. Case classes are similar to
algebraic datatypes but, unlike algebraic datatypes, new cases can be
modularly added. The reason why case classes are not a complete
solution to the Expression Problem is that they do not enforce
exhaustiveness of pattern matching for extensible operations. In other
words run-time pattern matching errors can happen when writing extensible code with case classes. So that full static type-safety is not ensured. Nevertheless case classes are a pragmatic approach, which is
widely used in practice. Therefore, modular parsing techniques may be
of value for extensible code using case classes.  The approach we
presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
all that the users need to do is to use case classes, instead of
standard OO classes in their code.
